Automating Verification of MPC Security

### 2023-06-03 popl24/minicat.tex
checkpoint

\section{The $\minicat$ Language}

The language $\minicat$ comprises a standard syntax of
boolean algebra and arithmetic expressions over integers. In addition, we allow reference
to \emph{variables} in expressions. \emph{Programs} are possibly empty
lists of assignments of expressions to variables.
$$
\begin{array}{rcl@{\qquad\qquad}r}
x,y,z &\in& \mathcal{V} & \textit{variables}\\
n &\in& \mathbb{Z} & \textit{integers}\\
b &::=& \true \mid \false & \textit{booleans}\\
v &::=& n \mid b & \textit{values} \\[2mm]
\mathit{unop} &::=& \mathrm{not} \mid - & \textit{unary operations} \\
\mathit{binop} &::=& \mathrm{and} \mid \mathrm{or} \mid + \mid * \mid \mod \mid\ =\ \mid & \textit{binary operations} \\[2mm]
e &::=& x \mid v \mid\, \mathit{unop}\, e \mid e\, \mathit{binop}\, e \mid  \eite{e}{e}{e}  & \textit{expressions}\\[2mm]
\prog &::=& \varnothing\quad \mid\quad \eassign{x}{e};\ \prog & \textit{programs}
\end{array}
$$
So for example, the following is a program- note that it assume the existence of an
\emph{input} variable $x$, which is undefined by the program itself:
$$
\eassign{y}{x + 1}; \eassign{z}{2 * y}
$$

For variable-free expressions in $\minicat$,
we can extend our interpretation of
boolean algebra $\cod{\cdot}$ to integer arithmetic in the obvious
manner, so for example $\cod{2 * 4} = 8$. However, to resolve
expressions that include variables, we need to define the semantics in
terms of references to \emph{memories}. We take the view that memories
map variables to values.
\begin{definition}[Memories]
A \emph{memory}, aka \emph{store} $\store$ is a finite (partial) mapping from variables to values.
The \emph{domain} of a memory is written $\dom(\store)$ and is the finite
set of variables on which the memory is defined. We write $\store\{ x \mapsto v\}$
to denote the memory $\store'$ such that $\store'(x) = v$ and otherwise
$\store'(y) = \store(y)$ for all $y \in (\dom(\store) - \{ x \}) $.
\end{definition}
Now we can modify our interpretation of expressions to take into account a given memory.
We write $\cod{\store, e}$ to mean the interpretation of $e$ where each variable $x$ in $e$
is replaced by $\store(x)$. In full detail, we define this as follows:
\begin{eqnarray*}
\cod{\store, x} &=& \store(x)\\
\cod{\store, \mathit{unop}\ e} &=& \cod{\mathit{unop}\ \cod{\store,e}}\\
\cod{\store, e_1\ \mathit{binop}\ e_2} &=& \cod{\cod{\store,e_1}\ \mathit{binop}\ \cod{\store,e_2}}\\
\cod{\store, \eite{e_1}{e_2}{e_3}} &=&
             \begin{cases}
                \cod{\store,e_2} & \text{if\ } \cod{\store,e_1}\\
                \cod{\store,e_3} & \text{otherwise}
             \end{cases}
\end{eqnarray*}

A critical point is that the semantics of $\minicat$ is defined not
just with respect to programs, but with respect to a pair of a memory
and a program $(\store, \prog)$, which we call
a \emph{configuration}. Together with the interpretation of
expressions that reference memories, we can obtain
an \emph{operational} semantics for configurations. This semantics
specifies how memories evolve during computation as a result of
assignments to variables in programs. The fundamental relation is
the \emph{reduction} aka \emph{evaluation} relation $\redx$,
which is a relation on configurations, that we define as follows:
$$
(\store, \eassign{x}{e};\prog) \redx (\extend{\store}{x}{\cod{\store,e}}, \prog)
$$

\subsection{Synchronous Distributed Protocols} With some minor syntactic
modifications to $\minicat$, we can model and reason about security in
protocols for distributed communications. We will call this modified
system $\fedcat$, alluding to the existence of a group of participants
we call a \emph{federation}. We restrict ourselves to boolean values
in $\fedcat$, which we will denote $1$ and $0$ for convenience.  We
assume the full set of the usual boolean operations including xor.

We identify participants by integers $n$, and assume that the
federation can be partitioned into two sets $H$ and $C$ where $H$ are
all \emph{honest} and $C$ are \emph{corrupt}. Then, we modify
variables $x$ to be of the form $\sx{n}{s}$ denoting a \emph{secret}
belonging to party $n$ and distinguished by a string $s$, or of
the form $\vx{n}{s}$ denoting a \emph{view} belonging to party $n$ and
distinguished by $s$. A view is a memory location that records a value
received by a party. So for example, in the following program, party 1
sends the negation of their secret $\ttt{foo}$ to party 2, with the
value recorded in party 2's $\ttt{bar}$ view:
$$
\eassign{\vx{2}{bar}}{\Dnot (\sx{1}{foo})}
$$ In $\fedcat$ we require that all assignments are to views, and in
any assignment $\eassign{\vx{n}{s}}{e}$ we require all variables that
occur in $e$ to belong to the same party. We will also assume that
secrets are defined in initial memories, that views are not, and that
views are never redefined (overwritten). This models communication
between parties through views, where each party can compute locally on
its own values, and ensures that the final memory after computation
contains a record of all secrets and views. We will use
\emph{protocol} as a synonym for program in $\fedcat$ given our intent
to model synchronous distributed computations.

 Although $\fedcat$ is a simple
language we can explore some surprisingly complex distributed
protocols with the introduction of randomized elements. In $\fedcat$
we accomplish this by introducing a new form of variables called
\emph{flips}, written $\fx{n}{s}$, which are owned by a party $n$ and
distinguished by string $s$. We will assume that flips can be either
$1$ or $0$ with equal probability.  So for example, in this
program, party 1 sends their flip $\ttt{foo}$ to party 2,
with the value recorded in party 2's $\ttt{bar}$ view:
$$
\eassign{\vx{2}{bar}}{\fx{1}{foo}}
$$
We don't change anything about the semantics of $\fedcat$
to model this randomization, rather we will assume that all
flip values are randomly assigned in initial memories. This is
the so-called ``random tape'' approach.

>>>>

\section{The $\minifed$ Language}

$$
\begin{array}{rcl@{\hspace{8mm}}r}
b &\in& \{ \etrue, \efalse \} \\
w &\in& \mathrm{String} \\ 
\cid &\in& \mathrm{Clients} \qquad  (\mathrm{Clients} \subset \mathbb{N} \text{\ in\ this\ presentation})\\[2mm]
\bop &\in& \{ \eand, \eor, \exor \} \\[2mm]
\be &::=& b \mid \flip{\cid}{w} \mid \secret{\cid}{w} \mid \view{\cid}{w} \mid \oracle{w} \mid & \textit{boolean expressions}\\
& &  \enot\ \be \mid \be\ \bop\ \be \mid \select{\be}{\be}{\be} \mid \OT{\be}{\be}{\be} \\[2mm]
\instr &::=& \eassign{\view{\cid}{w}}{\be} & \textit{view assignments} \\[2mm]
\prog &::=& \varnothing \mid \instr; \prog & \textit{protocols}
\end{array}
$$

\begin{definition}[Memories]
A \emph{memory}, aka \emph{store} $\store$ is a finite (partial) mapping from variables to values.
The \emph{domain} of a memory is written $\dom(\store)$ and is the finite
set of variables on which the memory is defined. We write $\store\{ x \mapsto v\}$
for  $x\not\in\dom(\store)$ to denote the memory $\store'$ such that $\store'(x) = v$ and otherwise
$\store'(y) = \store(y)$ for all $y \in \dom(\store)$.
\end{definition}

\begin{eqnarray*}
\lcod{\store, \etrue}{\cid} &=& \etrue\\
\lcod{\store, \efalse}{\cid} &=& \efalse\\
\lcod{\store, \flip{\cid}{w}}{\cid} &=& \store(\flip{\cid}{w})\\
\lcod{\store, \secret{\cid}{w}}{\cid} &=& \store(\secret{\cid}{w})\\
\lcod{\store, \view{\cid}{w}}{\cid} &=& \store(\view{\cid}{w})\\
\lcod{\store, \oracle{w}}{\cid} &=& \store(\oracle{w})\\
\lcod{\store, \enot\ \be}{\cid} &=& \cod{\enot\ \lcod{\store,\be}{\cid}}\\
\lcod{\store, \be_1\ \mathit{binop}\ \be_2}{\cid} &=&
    \cod{\lcod{\store,\be_1}{\cid}\ \mathit{binop}\ \lcod{\store,\be_2}{\cid}}\\
\lcod{\store, \select{\be_1}{\be_2}{\be_3}}{\cid} &=&
             \begin{cases}
                \lcod{\store,\be_2}{\cid} & \text{if\ } \lcod{\store,\be_1}{\cid}\\
                \lcod{\store,\be_3}{\cid} & \text{if\ } \neg\lcod{\store,\be_1}{\cid}
             \end{cases}%\\
%\lcod{\store, \OT{\be_1}{\be_2}{\be_3}}{\cid} &=&
%             \begin{cases}
%                \lcod{\store,\be_2}{\cid'} & \text{if\ } \lcod{\store,\be_1}{\cid}\\
%                \lcod{\store,\be_3}{\cid'} & \text{if\ } \neg\lcod{\store,\be_1}{\cid}
%             \end{cases}
\end{eqnarray*}

\begin{mathpar}
  (\store, \eassign{\view{\cid}{w}}{\be};\prog) \redx (\extend{\store}{\view{\cid}{w}}{\lcod{\store,\be}{\cid'}}, \prog)

  \inferrule
  {b = \text{if\ } \lcod{\store,\be_1}{\cid}  \text{\ then\ } \lcod{\store,\be_2}{\cid'} \text{\ else\ } \lcod{\store,\be_3}{\cid'}}
      {(\store, \eassign{\view{\cid}{w}}{\OT{\be_1}{\be_2}{\be_3}};\prog) \redx (\extend{\store}{\view{\cid}{w}}{b}, \prog)}
\end{mathpar}

### 2023-06-03 popl24/pmf.tex
checkpoint

Now we can define the probability distribution of a program $\prog$,
that we denote $\progd(\prog)$. Recall that $\fedcat$ is
deterministic, so a run is determined by the input secrets together
with random flip assignments. And since we constrain programs to not
overwrite memories, we can be sure that final memories resulting from
computation contain both a complete record of all initial secrets and
flips as well as views of communicated information. Thus a final
memory represents a run of the program, and since we consider and set
of \emph{input} secrets and flips to be equally likely, we consider
each memory in the set of final memories of a program to be equally
likely, which establishes the basic probability distribution for the
program. And since we don't care about flip values which are only
obversable insofar as they affect views, the probability distribution
we are really interested in is the marginal distribution of secrets
and views in this basic program distribution.
\begin{definition}
  Given progam $\prog$ with $\iov(\prog) = (S,V)$ and $\flips(\prog) = F$, define $M$ as:
  $$
  M \defeq \{ \store \mid (\dom(\store) = S \cup V \cup F) \wedge (\store_{S \cup F},\prog) \redxs (\store,\varnothing) \}
  $$
  Define also $\pdf{S \cup V \cup F}$ as the program's \emph{basic distribution} such that for all
  $\store \in \mems(S \cup V \cup F)$:
  $$
  \pdf{S \cup V \cup F}(\store) =
  \begin{cases}
    1 / |M| & \text{if}\ \store \in M\\
    0 & \text{otherwise}
  \end{cases}
  $$
  Then the \emph{program distribution of $\prog$}, denoted $\progd(\prog)$, is the
  marginal distribution of $S \cup V$ in $\prog$'s basic distribution:
  $$
  \progd(\prog) =  \margd{\pdf{S \cup V \cup F}}{S\cup V}
  $$
\end{definition}

>>>>

Now we can define the probability distribution of a program $\prog$,
that we denote $\progd(\prog)$. Recall that $\fedcat$ is
deterministic, so a run is determined by the input secrets together
with random flip assignments. And since we constrain programs to not
overwrite memories, we can be sure that final memories resulting from
computation contain both a complete record of all initial secrets and
flips as well as views of communicated information. Thus a final
memory represents a run of the program, and since we consider and set
of \emph{input} secrets and flips to be equally likely, we consider
each memory in the set of final memories of a program to be equally
likely, which establishes the basic probability distribution for the
program. And since we don't care about flip values which are only
obversable insofar as they affect views, the probability distribution
we are really interested in is the marginal distribution of secrets
and views in this basic program distribution.
\begin{definition}
  Given progam $\prog$ with $\iov(\prog) = (S,V)$ and $\flips(\prog) = F$, define $\stores$ as:
  $$
  \stores \defeq \{ \store \mid (\dom(\store) = S \cup V \cup F) \wedge (\store_{S \cup F},\prog) \redxs (\store,\varnothing) \}
  $$
  Define also $\pdf{S \cup V \cup F}$ as the program's \emph{basic distribution} such that for all
  $\store \in \mems(S \cup V \cup F)$:
  $$
  \pdf{S \cup V \cup F}(\store) =
  \begin{cases}
    1 / |\stores| & \text{if}\ \store \in \stores\\
    0 & \text{otherwise}
  \end{cases}
  $$
  Then the \emph{program distribution of $\prog$}, denoted $\progd(\prog)$, is the
  marginal distribution of $S \cup V$ in $\prog$'s basic distribution:
  $$
  \progd(\prog) =  \margd{\pdf{S \cup V \cup F}}{S\cup V}
  $$
\end{definition}
